---
title: "Homework 1"
author: "Apoorva Srinivasan"
date: "9/20/2018"
output: github_document
---

#PROBLEM 1

**This problem focuses on variable types and coercion.**



_The code chunk below creates a data frame of a random sample of size 10 between 0 and 5, a logical vector indicating whether elements of the sample are greater than 2, a character and a factor vector._

```{r hw1_problem1.1}
library(tidyverse)
set.seed(1)
problem1_df = tibble(
rsamp = runif(10, 0, 5),
vec_log = rsamp > 2,
vec_char = c("I", "hope", "I", "did", "okay", "for", "my", "first", "P8105", "homework"),

vec_fac = factor(c("heads", "tails", "heads", "heads", "tails", "heads", "tails", "tails", "heads", "tails"))


)



```



_Now, we're taking the mean of each variable in the data frame._

```{r hw1_problem1.2, warning = FALSE} 
mean(problem1_df$rsamp)
mean(problem1_df$vec_log)
mean(problem1_df$vec_char)
mean(problem1_df$vec_fac)
```

We observe that the random sample and logical vector return a mean of ~2.76 and ~0.6 respectively. However, we do not get any results for the factor and character variables. This is due to a numeric value attached to the first two vectors so mathematical operations on it is made possible. In the case of character and factor variable types, since they contain strings and do not posses a numerical value hence the mean cannot be found. 


_Onto converting variable types_
```{r hw1_problem1.3, WARNING = FALSE}
as.numeric(problem1_df$vec_log)
as.numeric(problem1_df$vec_char)
as.numeric(problem1_df$vec_fac)
```

*   On converting the logical variable to numeric type, R assigns '0' each FALSE condition i.e. for each value from the random sample less than 2 and a '1' for a TRUE condition i.e. every value which is greater than 2.
*   For the character variable type, the numeric conversion returned nothing since it could not assign a numeric value to random letters
*  In the conversion of factor to numerical type, R assigns the value '1' to heads and '2' to tails.

```{r hw1_problem1.4, warning = FALSE}
as.numeric(as.factor(problem1_df$vec_char))
as.numeric(as.character(problem1_df$vec_fac))
```




#PROBLEM 2
**This problem focuses on plotting and the use of inline R code.**

In the code chunk below, we've created a data frame comprised of a random sample 'x' and 'y' of size 1000 from a normal standard distribution and a logical, numeric and factor vector.
```{r}

library(tidyverse)
set.seed(1)

problem2_df = tibble(
 x = rnorm(1000),
 y = rnorm(1000),
vec_logical = c(x + y > 0),
vec_numerical = as.numeric(vec_logical),
vec_factor = as.factor(vec_logical)
)

```

*  The size of the dataset is `r length(problem2_df$x)`

*  The mean of x is `r mean(problem2_df$x)` and the median is `r median(problem2_df$x)`
*  The proportion of logical vector is `r sum(problem2_df$vec_logical) / length(problem2_df)`



Making a scatter plot using the logical variable.
```{r}

ggplot(problem2_df, aes(x = x, y = y, color = vec_logical)) + geom_point()
ggsave("logical_scatter_plot.pdf")
```



Making a scatter plot using the numerical variables.

```{r}
ggplot(problem2_df, aes(x = x, y = y, color = vec_numerical)) + geom_point()
```

Now, making the scatter plot using the factor variable

```{r}
ggplot(problem2_df, aes(x = x, y = y, color = vec_factor)) + geom_point()
```




