---
title: "Homework 1"
author: "Apoorva Srinivasan"
date: "9/20/2018"
output: github_document
---

#Problem 1

**This problem focuses on variable types and coercion.**


```{r hw1_problem1.1, include = FALSE}
##loading the library tidyverse but not displaying it in the output
 library(tidyverse)
```

_The code chunk below creates a data frame of a random sample of size 10, a logical vector indicating whether elements of the sample are greater than 2, a character and a factor vector._

```{r}
problem1_df = tibble(
  rsamp = runif(10, 0, 5),
  vec_log = rsamp > 2,
  vec_char = c("I", "hope", "I", "did", "okay", "for", "my", "first", "P8105", "homework"),
  vec_fac = factor(c("heads", "tails", "heads", "heads", "tails", "heads", "tails", "tails", "heads", "tails"))
)
```



_Now, we're taking the mean of each variable in the data frame._

```{r hw1_problem1.2, warning = FALSE} 
mean(problem1_df$rsamp)
mean(problem1_df$vec_log)
mean(problem1_df$vec_char)
mean(problem1_df$vec_fac)
```

We observe that the random sample and logical vector return a value of 1.935042 and 0.4 respectively. However, we do not get any results for the factor and character variables. This is due to a numeric value attached to the first two vectors so mathematical operations on it is made possible. In the case of character and factor variable types, since they contain strings and do not posses a numerical value hence the mean cannot be found. 

